Leadership across cultures and contexts: an exploratory analysis

### 2020-05-07 leadership_across_cultures_contexts.Rmd
Minor edits and code changes

To determine which leader qualities and functions were associated with evidence for high status, we fit a logistic elasticnet model of *High status* as a function of all quality and function variables, with the lasso penalty ($\alpha = 1$). The non-zero coefficients indicated that *Punishment* was a negative predictor, all other functions, including *Organize cooperation*, *Ritual functions*, and *Provide counsel/direction* were non-predictors, and *Wealthy*, *Strategic nepotism*, *Age*, *Experienced/accomplished*, and several others, were positive predictors. See Figure \@ref(fig:elastichighstatus).

```{r elastichighstatus, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *High Status*. A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *High status* in each text record. B: Non-zero coefficients of all quality and function variables that best predicted evidence for *High status*. Both models used the lasso penality ($\\alpha=1$), with $\\lambda_{1se}$ chosen by cross-validation (1 SE from \\lambda_{min}). Red values are positive predictors; purple values are negative predictors. Coefficents = 0 not displayed. X-axes are odds ratios on a log 10 scale."}

@@

```{r elasticcoercive, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *Coercive authority*. A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *Coercive authority* in each text record ($\\alpha = 0.5$). B: Non-zero coefficients of all quality and function variables that best predicted evidence for *Coercive authority* ($\\alpha=1$). Both models chose $\\lambda_{1se}$ by cross-validation (1 SE from \\lambda_{min}). Red values are positive predictors; purple values are negative predictors. Coefficents = 0 not displayed. X-axes are odds ratios on a log 10 scale."}

@@

We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures. We found evidence that leaders function to resolve conflict, organization cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligence, and experienced and accomplished in about 80% or more of cultures. See Figure \@ref(fig:varsupport). Leaders benefitted materially, reproductively, and socially in 50% or more of cultures. See Figure \@ref(varsupportcb). Some common leader dimensions also did not vary by continental region, subsistence strategy, group context, or leader sex. These included qualities, such as being knowledgeable and intelligence, wealthly, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group. These dimensions are therefore candidates for universal dimensions of human leadership. By "universal", we do not mean that all leaders have all these traits. Instead, we mean that these traits are found in some leaders across all cultures and contexts.

Many leadership dimensions did vary across cultures and contexts, however, including functions such as organizing cooperation, providing counsel and direction and prosocial investment, and qualities such as high status and aggressiveness. See Figure \@ref(fig:heatmapdrop1)

@@

We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is much greater than the number of cases (i.e., $p > > n$). Lasso regression has a penalty or shrinkage term, $\lambda$: larger values of $\lambda$ retain more variables in the model, at the risk of over-fitting, whereas smaller values of $\lambda$ decrease the number of predictors retained in the model, at the risk of under-fitting. The optimal value of $\lambda$ can be estimated with cross-validation [@tibshirani_regression_1996; @zou_regularization_2005]. We used the glmnet package [@friedman_regularization_2010] to fit a lasso regression model to determine which words best predicted evidence for high status as a leader quality. We used tenfold cross-validation to choose the $\lambda$ parameter that minimized deviance. We then plotted all the non-zero coefficients (Figure \@ref(fig:highstatusplot)).

```{r highstatusplot, echo=F, message=F, fig.width=8, fig.height=6, fig.align='center', fig.cap="Lasso logistic regression coefficients, which indicate the words whose frequencies in each text record best predicted evidence for the *high status* leader quality in each text record. Coefficients estimated with glmnet package using cross-validation to determine the optimal value of lambda [@friedman_regularization_2010]. Red values are positive predictors; purple values are negative predictors. X-axis is on a log 10 scale."}
# highstatus_plot + scale_x_log10()
```

>>>>

To determine which leader qualities and functions were associated with evidence for high status, we fit a logistic elasticnet model of *High status* as a function of all quality and function variables, with the lasso penalty ($\alpha = 1$). The non-zero coefficients indicated that *Punishment* was a negative predictor, all other functions, including *Organize cooperation*, *Ritual functions*, and *Provide counsel/direction* were non-predictors, and *Wealthy*, *Strategic nepotism*, *Age*, *Experienced/accomplished*, and several others, were positive predictors. See Figure \@ref(fig:highstatusplot).

```{r highstatusplot, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *High Status*. A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *High status* in each text record. B: Non-zero coefficients of all quality and function variables that best predicted evidence for *High status*. Both models used the lasso penality ($\\alpha=1$), with $\\lambda_{1se}$ chosen by cross-validation (1 SE from $\\lambda_{min}$). Red values are positive predictors; purple values are negative predictors. Variables with coefficents = 0 not displayed. X-axes are odds ratios on a log 10 scale."}

@@

```{r elasticcoercive, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *Coercive authority*. A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *Coercive authority* in each text record. B: Non-zero coefficients of all quality and function variables that best predicted evidence for *Coercive authority*. Both models used the lasso penalty ($\\alpha=1$) and chose $\\lambda_{1se}$ by cross-validation (1 SE from $\\lambda_{min}$). Red values are positive predictors; purple values are negative predictors. Variables with coefficents = 0 not displayed. X-axes are odds ratios on a log 10 scale."}

@@

We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures. Leaders function to resolve conflict, organization cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligence, and experienced and accomplished in about 80% or more of cultures. See Figure \@ref(fig:varsupport). Leaders benefitted materially, reproductively, and socially in 50% or more of cultures. See Figure \@ref(fig:varsupportcb). Common leader dimensions that did not vary by continental region, subsistence strategy, group context, or leader sex included qualities, such as being knowledgeable and intelligence, wealthly, strategically nepotistic, and generous; and functions, such as resolving conflict and representing the group. These dimensions are therefore candidates for universal dimensions of human leadership. By "universal", we do not mean that all leaders have all these traits. Instead, we mean that there is a fairly high level of evidence for these traits across cultures that does not vary by continental reagion, subsistence strategy, group context, or leader sex. See Table \@ref(tab:leaderuniversals).

Evidence for many leadership dimensions did vary across cultures and contexts, however, including functions such as organizing cooperation, providing counsel and direction and prosocial investment; and qualities such as high status and aggressiveness. See Figure \@ref(fig:heatmapdrop1).

@@

We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is large relative to the number of cases. Lasso regression has a penalty or shrinkage term, $\lambda$: larger values of $\lambda$ retain more variables in the model, at the risk of over-fitting, whereas smaller values of $\lambda$ decrease the number of predictors retained in the model, at the risk of under-fitting. The optimal value of $\lambda$ can be estimated with cross-validation [@tibshirani_regression_1996; @zou_regularization_2005]. We used the glmnet package [@friedman_regularization_2010] to fit a lasso regression model to determine which words best predicted evidence for high status as a leader quality. We used tenfold cross-validation to choose a $\lambda$ parameter that was within in standard error of the value that minimized the deviance. We then plotted all the non-zero coefficients (Figure \@ref(fig:highstatusplot)).

```{r highstatusplot2, echo=F, message=F, fig.width=8, fig.height=6, fig.align='center', fig.cap="Lasso logistic regression coefficients, which indicate the words whose frequencies in each text record best predicted evidence for the *high status* leader quality in each text record. Coefficients estimated with glmnet package using cross-validation to determine the optimal value of lambda. Red values are positive predictors; purple values are negative predictors. X-axis is on a log 10 scale."}
# highstatus_plot + scale_x_log10()
```
